Coming to Germany: Pre-migration language skills, Job search and Labor market outcomes

### 2018-08-31 Paper 2.tex
Update Paper 2.tex

Language skills are the key component of an immigrant's human capital throughout his or her migration history,  (Isphording, 2014). Empirical studies predominantly find positive effects of language skills on labor market success of immigrants throughout a range of countries and languages, such as English in the UK (Dustmann and Fabbri, 2003), the US (Bleakley and Chin, 2004), Australia (Chiswick and Miller, 1995); German in Germany (Dustmann and van Soest, 2002, 2002; Isphording et al., 2014), Spanish and Catalan in Spain (Burda and Swedberg, 2012; Isphording, 2013). Typical issues emerged in measuring the labor market effects of language skills are measurement errors in survey data, reverse causality and omitted variable bias (Yao and Ours, 2015). Therefore, most recent empirical studies rely on an instrumental variable (IV) to overcome these problems such as using leads and lags of language skills and fathers' education (Dustmann and van Soest, 2001, 2002), age at arrival (Bleakley and Chin, 2004, 2010; Budra and Swedberg, 2012; Di Paolo and Raymond, 2012; Miranda and Zhu, 2013a, 2013b), or linguistic distance (Isphording, 2013; Isphording et al., 2014). IV estimates are generally higher than OLS estimates, which shows that the downward bias from measurement errors dominates the upward bias from individual unobserved characteristics and reverse causality (Yan and Ours, 2015).

Although IV estimations provide a robust estimate of the labor market effect of language skills, they give relatively little insightful information about the mechanism through which such an effect exists (Isphording, 2014). Some authors attempt to document some major channels through which language skills affect the earnings of immigrants, for instance, through complimenting pre--migration schooling and experience (Chiswick and Miller, 2003), occupation level (Berman et al., 2003), increased education (Bleakley and Chin, 2004; Isphording and Sinning, 2012), and occupational choice (Chiswick and Miller, 2010).

@@

\subsection{The job search behavior of immigrants}
Like natives, immigrants also use a variety of methods to seek for jobs. These methods include the use of public and private employment agencies, responding to job advertisements on public media (newspapers or the Internet), or relying on their network of family and friends (Battu et al., 2011). The point of departure in previous literature is the contemporaneous language skills, therefore, they do not discuss the job search behaviors of immigrants. But as a most important and visible human capital component in the migration history, language skills, in particular, pre--migration language skills obviously have an impact on the choice of job search methods. And these job search methods, in turn, do matter in subsequent labor market outcomes (Battu et al., 2011).

>>>>

Language skills are the key component of an immigrant's human capital throughout his or her migration history  (Isphording, 2014). Empirical studies predominantly find positive effects of language skills on labor market success of immigrants throughout a range of countries and languages, such as English in the UK (Dustmann and Fabbri, 2003), the US (Bleakley and Chin, 2004), Australia (Chiswick and Miller, 1995); German in Germany (Dustmann and van Soest, 2002, 2002; Isphording et al., 2014), Spanish and Catalan in Spain (Burda and Swedberg, 2012; Isphording, 2013). Typical issues emerged in measuring the labor market effects of language skills are measurement errors in survey data, reverse causality and omitted variable bias (Yao and Ours, 2015). Therefore, most recent empirical studies rely on instrumental variables (IVs) to overcome these problems such as using leads and lags of language skills and fathers' education (Dustmann and van Soest, 2001, 2002), age at arrival (Bleakley and Chin, 2004, 2010; Budra and Swedberg, 2012; Di Paolo and Raymond, 2012; Miranda and Zhu, 2013a, 2013b), or linguistic distance (Isphording, 2013; Isphording et al., 2014). IV estimates are generally higher than OLS estimates, which shows that the downward bias from measurement errors dominates the upward bias from individual unobserved characteristics and reverse causality (Yan and Ours, 2015).

Although IV estimations provide a robust estimate of the labor market effect of language skills, they give relatively little insightful information about the mechanism through which such an effect exists (Isphording, 2014). Some authors attempt to document some major channels through which language skills affect the earnings of immigrants, for instance, through complimenting pre--migration schooling and experience (Chiswick and Miller, 2003), occupation level (Berman et al., 2003), increased education (Bleakley and Chin, 2004; Isphording and Sinning, 2012), and occupational choice (Chiswick and Miller, 2010). However, it is empirically not clear whether the pre--migration language skills also work through these channels like contemporaneous language skills.

@@

\subsection{The job search behavior of immigrants}

Like natives, immigrants also use a variety of methods to seek for jobs. These methods include the use of public and private employment agencies, responding to job advertisements on public media (newspapers or the Internet), or relying on their network of family and friends (Battu et al., 2011). Job search theory predicts that people choose their search intensity and search methods by comparing the returns and costs of different options (Holzer, 1988). Hence, as a most important and visible human capital component in the migration history, pre--migration language skills obviously have an impact on the choice of job search methods. And these job search methods, in turn, do matter in subsequent labor market outcomes (Battu et al., 2011).

Most of studies find a positive impact of using personal social networks on the probability of obtaining jobs, however, empirical results of the wage effects are inconclusive (Alaverdyan, 2017). Dustmann et al., (2011) find a positive effect on job matching quality of using informal social networks. Lancee (2016) also finds a positive wage effects of using referrals for migrants in Germany, but the result holds only for highly educated immigrants with good German language skills. On the contrary, Battu et al., 2011 and Carlsson et al., 2014 conclude that using social networks is an effective method to obtain job offers, but it is associated with lower wages.

Cappellari and Tatsiramos (2015) focus on the network quality as a key factor to explain the effectiveness of social contacts on labor market performance. They find that a higher number of employed friends within the closest tie increases both the job finding rate and wages, but only for high-skilled workers. Facchini et al., (2015) use the German Social-Economic Panel (GSOEP) 1996 -- 2011 to find that immigrants who have German friends in their friendship ties tend to have faster cultural assimilation than immigrants who do not. While they argue that labor market integration and education are important determinants of forming friendship ties with natives and not the other way around, there seems to be a strong inter-relation between labor market success and the quality of social networks.

% The point of departure in previous literature is the contemporaneous language skills, therefore, they do not discuss the job search behaviors of immigrants in measuring the labor market effects of language skills. But as
